---
title: "Develop ea_homogeneous_areas"
author: "Anders Kolstad"
date: "2023-01-17"
output: html_document
---

Develop ea_homogeneous_areas to include delim2

```{r}
library(devtools)
load_all()
library(tmap)
library(stars)
library(dplyr)
library(stats)
library(ggplot2)
```

Get test data

```{r}
data("ex_polygons")
data("accounting_area")
data("ex_raster")
```

```{r}
ea_homogeneous_area_dev <- function(delineation1,
                                delineation2,
                                outline,
                                keep1 = NULL,
                                keep2 = NULL,
                                groups){
  if("stars" %in% class(delineation1)){
    
    if(max(dim(delineation1)) > 10000) warning("You have inputted a large raster file. Vectorisation might take a long time.")
    if(missing(groups)) warning("No groups argument provided. Please check output if this is really what you want.")
    
    groups_int <- enquo(groups)
    myArea <- sf::st_as_sf(delineation1, 
                           as_points = FALSE, 
                           merge = FALSE) %>%
      group_by(!!groups_int) %>%
      summarise(geometry = sf::st_union(geometry)) %>%
      ungroup() %>%
      st_cast(to = "POLYGON")
  }
  if("sf" %in% class(delineation1)){
    myArea <- delineation1 %>%
      select(all_of(keep1))
  }
  if(missing(delineation2)) {
    return(myArea)
  } else{
    if("sf" %in% class(delineation2)){
      myArea <- delineation2 %>%
        select(all_of(keep2)) %>%
        st_intersection(myArea) %>%
        st_collection_extract() %>%
        st_cast(to = "POLYGON")
      if(missing(outline)){
        return(myArea)
      } else {
        if("sf" %in% class(outline)){
          myArea_outline <- myArea %>%
            st_intersection(outline)
          return(myArea_outline)
        } else stop("Outline is not an sf object")
      }
    } else stop("Delineation2 is not an sf object")
  }
}
```

# Example 1: two sf objects

```{r}
ex1 <- ea_homogeneous_area_dev(ex_polygons,
                                accounting_area,
                                keep1 = "condition_variable_2",
                                keep2 = "name")
```

All geometries are valid (closed)

```{r}
table(st_is_valid(ex1))
```

All geometries are converted to polygons, including points and lines that might be produced as part of the intersection of overlapping geometries. See [here](https://github.com/r-spatial/sf/issues/1484).

```{r}
unique(st_geometry_type(ex1))
```

Only the names given by `keep1` and `keep2` are retained:

```{r}
names(ex1)
```

```{r}
tm_shape(ex1)+
  tm_polygons(col = "name")
```

# Example 2: One raster and one sf object

```{r}
ex2 <- ea_homogeneous_area_dev(ex_raster,
                                accounting_area,
                                keep2 = "name",
                               groups = values)


```

Calculate the area of the new polygons

```{r}
ex2$area <- units::drop_units(sf::st_area(ex2))
hist(ex2$area)

```

Visualise where the small polygons are. They should be produced at the boarder between intersecting polygons

```{r}
plot(ex2$geometry)
plot(ex2$geometry[ex2$area<1000000],
     col="black", add=T)

```

```{r}
tm_shape(ex2)+
  tm_polygons(col = "values")+
  tm_shape(ex2[ex2$name == "Enebakk",])+
  tm_polygons(col = "blue", border.col = "black")

```

## Example 2.1: Test on bigger data

```{r}
HIA <- readRDS("P:/41201785_okologisk_tilstand_2022_2023/data/infrastrukturindeks/homogeneous_impact_areas.rds")
regions <- sf::read_sf("../../ecosystemCondition/data/regions.shp")
```

```{r}
st_crs(HIA) == st_crs(regions)
```

```{r}
regions <- sf::st_transform(regions, 32633) 
```

```{r, eval=F}
ex3 <- ea_homogeneous_area(HIA,
                           regions,
                           keep1 = "infrastructureIndex",
                           keep2 = "region")
```

Runtime was about 30 seconds.

```{r, eval=F}
dim(ex3); dim(HIA); dim(regions)
```

[1] 32402 3

[1] 31869 2

[1] 5 3


# Example 3: Ex2 + outline
```{r}
enebakk <- accounting_area[accounting_area$name == "Enebakk",]
dim(enebakk)
```


```{r}
ex3 <- ea_homogeneous_area_dev(ex_raster,
                                accounting_area,
                               enebakk,
                                keep2 = "name",
                               groups = values)
```

```{r}
tm_shape(ex3)+
  tm_polygons(col = "values")
```

