---
title: "Test ea_spread"
author: "Anders Kolstad"
date: "2023-01-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Testing and developing ea_spread


```{r}
ea_spread_dev <- function(indicator_data,
                      indicator,
                      regions,
                      groups,
                      threshold = 1,
                      tally = FALSE,
                      simple_join = FALSE){
  ID <- SHAPE <-area <- indicator_NA <- meanIndicatorValue <- NULL
  if("sf" %in% class(indicator_data) & "sf" %in% class(regions)){
      # get the intersections
    st_agr(indicator_data) <- "constant"
    st_agr(regions) <- "constant"
    indicator_split <- sf::st_intersection(indicator_data, regions)
      # calculate area of the intersections
    indicator_split$area <- drop_units(sf::st_area(indicator_split))
    if(simple_join == TRUE) {
      return(indicator_split)
    }
  } else stop("The input data is not in a supported format. Both indicator_data and regions need to be sf objects.")
      # Calculate the area weighted mean indicator values for each region
  
  if(missing(indicator)) stop("Indicator column is not defined")
  if(missing(groups)) stop("Group is not defined")
  
  groups_int <- enquo(groups)
  indicator_int <- enquo(indicator)
  
  if(tally == FALSE){
  myWeightedMeans <- indicator_split %>%
      group_by(!!groups_int) %>%
      mutate(n = n(),
             indicator_NA = ifelse(n >=threshold, !!indicator_int, NA))%>%
      summarise(meanIndicatorValue =
                  stats::weighted.mean(x = indicator_NA,
                                       w = area)) %>%
      mutate("ID" = !!groups_int) %>%
      as.data.frame() %>%
      select(-SHAPE)

    # paste these new values into the regions data set
  regions <- regions %>%
    mutate("ID" = !!groups_int) %>%
    left_join(myWeightedMeans, by = "ID") %>%
    select(ID, meanIndicatorValue)
  return(regions)
  } else {
    tally <- as.data.frame(indicator_split) %>%
      group_by(!!groups_int) %>%
      summarise(data_points = n(),
                area = sum(area))
    return(tally)
  }
}
```

```{r}
data("ex_polygons")
data("ex_raster")
```

## Example 1: Aggregate and spread to raster
Example using a raster to define homogeneous area classes.
Zooming in on the example data

```{r}
ex_raster_zoom <- ex_raster[,5:6, 28:29]
ex_polygons_zoom <- sf::st_crop(ex_polygons, ex_raster_zoom)
```

Scale the indicator
```{r}
ex_polygons_zoom$indicator <- ea_normalise(ex_polygons_zoom,
"condition_variable_2",
upper_reference_level = 7)

```

Tweak the data slightly for exagerated effect
```{r}
ex_polygons_zoom$indicator[2:6] <- 1
```

Process the `ex_raster_zoom` and define homogeneous areas based on the cell values.
```{r}
myRegions <- ea_homogeneous_area(ex_raster_zoom)
```

Now use the function
```{r}
out1 <- ea_spread_dev(indicator_data = ex_polygons_zoom,
                 indicator = indicator,
                 regions = myRegions,
                 groups = values)
```

... and plot the results
```{r}
plot(out1[,2])
```

## Example 2: Simple tally

```{r}
ea_spread_dev(indicator_data = ex_polygons_zoom,
                 indicator = indicator,
                 regions = myRegions,
                 groups = values,
                 tally = T)
```

### Simple join
```{r}
out_3 <- ea_spread_dev(indicator_data = ex_polygons_zoom,
                 indicator = indicator,
                 regions = myRegions,
                 groups = values,
                 simple_join = T)

```

<!-- Now find a way to get the area weighted indicator values per region. -->
