---
title: "Aggregate indicators to regions"
output: 
  bookdown::html_document2:
    toc: yes
    toc_depth: 2
    number_sections: FALSE
pkgdown:
  as_is: true
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(eaTools)
library(tmap)
library(stars)
library(dplyr)

```

# Introduction
We are often in a position where we need to aggregate indicators to certain regions within a larger accounting area. This process is a little different depending on the spatial coverage of the indicator and whether the regions are defined in a raster or with polygons. 

```{r}
data("ex_polygons")
data("accounting_area")
data("ex_raster")
indicator <- ex_polygons
#regions <- accounting_area
regions <- ex_raster
```

```{r}
tm_shape(ex_raster)+
  tm_raster(alpha = .3,
            style = "cat")+
  tm_shape(regions)+
  tm_polygons(alpha = .3)+
  tm_shape(indicator)+
  tm_polygons()+
  tm_layout(legend.outside = T)
```

Case 1: Aggregate the indicator values from the small polygons for each of the three classes in the raster layer.

We can vectorize the raster
```{r}
p = st_as_sf(regions, as_points = FALSE, merge = TRUE)
```

```{r}
tm_shape(p)+
  tm_polygons(col = "values")+
  tm_shape(indicator)+
  tm_polygons()+
  tm_layout(legend.outside = T)
```

Pull values (categorical class 1 -3) from the vectorized raster over to to the small polygons 
```{r}
p2 <- st_intersection(indicator, p)
```

```{r}
tm_shape(p2)+
  tm_polygons(col = "condition_variable_2")+
  tm_shape(indicator)+
  tm_polygons()+
  tm_layout(legend.outside = T)
```
 
summarise the indicator per category
```{r}
p3 <- p2 %>%
  group_by(values) %>%
  dplyr::summarise(meanIndicator = mean(condition_variable_2))
```

and pull these values back to the vectorized grid
```{r}
p$meanIndicator <- p3$meanIndicator[match(p$values, p3$values)]
```

```{r}
tm_shape(p)+
  tm_polygons(col = "meanIndicator")+
  tm_shape(indicator)+
  tm_polygons()+
  tm_layout(legend.outside = T)
```

We can rasterize the polygons
```{r}
indicator_raster <- stars::st_rasterize(indicator,
                              template = regions,
                              #align = T,
                              options = c("ALL_TOUCHED=TRUE"))
```

```{r}
tm_shape(indicator_raster[3,,])+
  tm_raster()+
  tm_shape(indicator)+
  tm_polygons()
```


```{r}
summary(indicator$condition_variable_2)
```

