---
title: "Aggregate indicators to regions"
output: 
  bookdown::html_document2:
    toc: yes
    toc_depth: 2
    number_sections: FALSE
pkgdown:
  as_is: true
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, warning=F, message=F}
library(eaTools)
library(tmap)
library(stars)
library(dplyr)
library(stats)
library(ggplot2)
```

# Introduction

We are often in a position where we need to aggregate indicators to certain regions within a larger accounting area. This process is a little different depending on the spatial coverage of the indicator and whether the regions are defined in a raster or with polygons.

We will use some example data sets included in the package.

```{r}
data("ex_polygons")
data("accounting_area")
data("ex_raster")
```

```{r example-data, fig.cap="Showing the three example datasets on top of eachother. The raster layer (ex_raster) have cell values 1, 2 or 3, as shown in the legend. The five large polygons is the accounting_area layer, and the smaller polygins, visible as specks, is teh ex_polygons layer."}
tm_shape(ex_raster)+
  tm_raster(alpha = 1,
            style = "cat")+
  tm_shape(accounting_area)+
  tm_polygons(alpha = .6)+
  tm_shape(ex_polygons)+
  tm_polygons(border.col = "black")+
  tm_layout(legend.outside = T)+
  tm_scale_bar(position = c("right", "bottom"), width = 0.3)

```

# Case 1: Aggregate indicator values from small polygons for each of the three classes in the raster layer, and *spread them* to cover the entire raster layer.

In this example we are imagining that `ex_polygons`, and the column `condition_variable_2`, contains a condition variable, and that `ex_raster` denotes homogeneous areas. We aknowledge a spatial bias in the sampling of the dummy condition variable. However, in this hypothetical example, we think that the variable/indicator can be said to be representative of the homogeneous area class that it belongs to. We therefore allow ourselves to extrapolate, or *spread*, the mean indicator values within a given homogeneous area class to the other grid cells of the same class.

To be able to see what we are doing, we will zoom in on the data above.

```{r, warning=F}
ex_raster_zoom <- ex_raster[,5:6, 28:29]
ex_polygons_zoom <- sf::st_crop(ex_polygons, ex_raster_zoom)
```

First we would scale the variable, and make it into an indicator using `ea_normalise()` (See `vignette("Normalise condition variable")`).

```{r}
ex_polygons_zoom$indicator <- ea_normalise(ex_polygons_zoom,
                                           "condition_variable_2",
                                           upper_reference_level = 7)
```

```{r example-indicator-hist, fig.cap="Histogram of the example indicator values.", warning=F, message=F}
ggplot(ex_polygons_zoom, aes(x = indicator))+
  geom_histogram(fill = "grey",
                 colour = "black",
                 size = 1.2)+
  theme_bw(base_size = 12)
```

```{r exampla-data-zoom, fig.cap="Example data produced by zoomin in on four grid cells in the above figure."}
tm_shape(ex_raster_zoom)+
  tm_raster(style="cat")+
  tm_shape(ex_polygons_zoom)+
  tm_polygons(col = "indicator",
              palette = "RdYlGn",
              border.col = "black")+
  tm_layout(legend.outside = T)+
  tm_scale_bar(position = c("left", "bottom"), width = 0.3)
```

I will manipulate the data set slightly, just to make some contrasts between the three homogeneous area classes more visible.

```{r}
ex_polygons_zoom$indicator[2:6] <- 1
```

We now want to calculate the area weighted mean for each homogeneous area class, i.e. the three different coloured grid cells in Fig. \@ref(fig:exampla-data-zoom).

First we process the `ex_raster_zoom` and define *homogeneous areas* based on the cell values.

```{r}
myRegions <- ea_homogeneous_area(ex_raster_zoom)
```

Then we aggregate the indicator values from ex_polygons_zoom for each homogeneous area class, and spread the mean indicator values across all the grid cells. Even grid cells that originally had no indicator values associated with it will be populated with the area weighted mean of its homogeneous area class. Lets to this step with and without defining a threshold for how many data points we need before calculating an average.

```{r}
myIndicator <- ea_spread(indicator_data = ex_polygons_zoom,
                         indicator = indicator,
                         regions = myRegions,
                         groups = values,
                         threshold = 5)
```

```{r}
myIndicator_noThreshold <- ea_spread(indicator_data = ex_polygons_zoom,
                         indicator = indicator,
                         regions = myRegions,
                         groups = values)
```

```{r easpread-out, fig.cap="Example where the area weighted averages of the indicator values in the small polygons are spread over the raster cells with each homogeneous area class. In the left example, only homogeneous area class with more than 5 data points get indicator values."}
tmap_arrange(
  tm_shape(myIndicator)+
    tm_polygons(col = "meanIndicatorValue",
                title="Indicator value",
                palette = "RdYlGn",
                breaks = seq(0,1,0.2))+
    tm_shape(ex_polygons_zoom)+
    tm_polygons(col = "indicator",
                legend.show = F,
                palette = "RdYlGn",
                breaks = seq(0,1,0.2),
                border.col = "black")+
    tm_layout(legend.position = c('left', 'top'),
              title = "Threshold = 5 data points")
  ,
  tm_shape(myIndicator_noThreshold)+
    tm_polygons(col = "meanIndicatorValue",
                title="Indicator value",
                palette = "RdYlGn",
                breaks = seq(0,1,0.2))+
    tm_shape(ex_polygons_zoom)+
    tm_polygons(col = "indicator",
                legend.show = F,
                palette = "RdYlGn",
                breaks = seq(0,1,0.2),
                border.col = "black")+
    tm_layout(legend.position = c('left', 'top'),
              title = "No threshold")
)
```

In the map above, the color of the cells are in a way the average color of the polygons that overlay these cells. Note that the bottom left cell now still has an indicator value even though it didn't have any data to start with.

If we look at the right hand figure, homogeneous area class 3 is in better condition, but this indicator value is based on very few data points. If we assign a threshold above the number of available data points, the function returns NA.

We can then rasterise this back again, because a raster format is more useful for subsequent analyses, such as aggregating indicators. Here we can use an existing function from the `sf` package.

```{r}
newIndicator_raster <- stars::st_rasterize(myIndicator, ex_raster_zoom)
```
