---
title: "Normalise condition variable"
output: 
  bookdown::html_document2:
    toc: yes
    toc_depth: 2
    number_sections: FALSE
pkgdown:
  as_is: true
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(eaTools)
```

# Introduction

This article explains the use the `ea_normalise()`. Use this function to normalise (also called rescale) condition variables based on defined reference levels (also called reference values).

Let's get some example data.

```{r}
data("ex_polygons")
class(ex_polygons)
```

This is an sf object with `r nrow(ex_polygons)` rows (i.e. polygons). The column `condition_variable_2` contains a dummy variable which we will take to be in raw units (e.g. kg or km).

```{r}
plot(ex_polygons$condition_variable_2)
```

We want to normalise this variable to become bound by 0 and 1.

# Plotting the results

It is helpful to visualise the transformation to make sure the function is doing what you expect of it (Fig. \@ref(fig:first-vis)).

```{r first-vis, fig.cap="Visualising the results of normalising and linear scaling of condition_variable_2 using an defined upper reference level and a default lower reference level of zero"}
eaTools::ea_normalise(data = ex_polygons,
             vector = "condition_variable_2",
             upper_reference_level = 7,
             plot = T)
```

This is in a way the default method with linear scaling between the upper and lower reference levels, and indicator values truncated above 1 and below 0. We can also add a break point (Fig. \@ref(fig:break-point)). The break point is the value of the condition variable which will become scaled to 0.6 (the threshold used to define *good ecological condition* in some contexts).

```{r break-point, fig.cap="Visualising the results of normalising and linear break-point scaling of condition_variable_2 using an defined upper reference level and a default lower reference level of zero"}
eaTools::ea_normalise(data = ex_polygons,
             vector = "condition_variable_2",
             upper_reference_level = 7,
             break_point = 5,
             plot = T)

```

We can do a two-sided normalisation by defining an optimum value (Fig. \@ref(fig:two-sided)). Note that this is non-compatible with defining a break-point.

```{r two-sided, fig.cap="Visualising the results of doing a two-sided normalising of condition_variable_2 using an defined upper reference level and a default lower reference level of zero"}
eaTools::ea_normalise(data = ex_polygons,
             vector = "condition_variable_2",
             upper_reference_level = 7,
             optimum = 5,
             plot = T)

```

Finally we also have the option of doing scaling the condition variable using a sigmoid function (Fig. \@ref(fig:sig)), and a positive (Fig. \@ref(fig:concave)) or negative (Fig. \@ref(fig:con)) exponential function. The latter two are non-compatible with a defined break-point, but the first one is.

```{r sig, fig.cap="Visualising the results of doing a normalising and sigmoid transformation of condition_variable_2 using an defined upper reference level and a default lower reference level of zero"}
eaTools::ea_normalise(data = ex_polygons,
             vector = "condition_variable_2",
             upper_reference_level = 7,
             scaling_function = "sigmoid",
             plot = T)
```

```{r con, fig.cap="Visualising the results of doing a normalising and negative exponensial transformation of condition_variable_2 using an defined upper reference level and a default lower reference level of zero"}
eaTools::ea_normalise(data = ex_polygons,
             vector = "condition_variable_2",
             upper_reference_level = 7,
             scaling_function = "exponential convex",
             plot = T)
```

```{r concave, fig.cap="Visualising the results of doing a normalising and positive exponensial transformation of condition_variable_2 using an defined upper reference level and a default lower reference level of zero"}
eaTools::ea_normalise(data = ex_polygons,
             vector = "condition_variable_2",
             upper_reference_level = 7,
             scaling_function = "exponential concave",
             plot = T)

```

## Export values

When you've found the correct settings for your normalisation, you can export the new values like this

```{r}
ex_polygons$condition_indicator_2 <- eaTools::ea_normalise(data = ex_polygons,
             vector = "condition_variable_2",
             upper_reference_level = 7)


```
